Refinements for Open Automata

### 2023-09-08 2021-Refinement/paper/refinementarticleV2.tex
remove highlight + minors errors

 \begin{itemize}
 \item   
For any open transition $OT$ in ${T}_1$:
$
     \OTbase{s \OTarrow{\alpha} s'}
         {
           \beta_j^{j\in J'}}{g_{OT}}{\psi_{OT}}
$
 there exists an indexed set of  open transitions $OT_x^{x\in X} \subseteq {T}_2$:
$
%    \left( fresh \ \set{\alpha_i}, \set{b_j}, v_x.\ \
    \OTbase{t \OTarrow{\alpha_x} t_x}
         {
           \beta_{j x}^{j\in J_{x}}}{ g_{OT_x}}{\psi_{OT_x}}   
%         \right)
$
%\end{minipage}
%\hspace{2mm}
%\begin{minipage}{0.35\linewidth}
%\vspace{-2em}
%{	\includegraphics[width=1.02\linewidth]{XFIG/Bisim}}
%\end{minipage}
 such that  $\forall x.\, J'=J_{x}$ and there exists \hl{ $Pred_{s',t_x}$ such that} $\mathcal{R}(s',t_x)  
 $ holds \RAB{On n'a pas  de Pred dans la def}
 and  
\begin{multline*}
 \mathcal{R}(s,t) \land g_{OT}\implies\\
 \displaystyle{\bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \land g_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \mathcal{R}\left(s',t_x\right)\psubst{\psi_{OT}\!\uplus\!\psi_{OT_x}}\right)}
\end{multline*}
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.
%\bigskip
%
% $\Pred_{s,t} \land \Pred_{OT}\implies\bigvee_{x\in X} \Pred_x$
%and\\
%%\hspace{1cm}
% $ \forall{x\in X}. \Pred_x \land \Pred_{s,t} \land \Pred_{OT} \Rightarrow
%   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
%     \land \alpha\!=\!\alpha_x \land  
%     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)$
%     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
%     %     \right)$.

@@

Similarly to FH-bisimulation~\cite{AMHEEMA:2023} we are interested  in finding simulation relations between configurations of two OA that contain variables and holes. When dealing with open systems it is common to define simulation in terms of a simulation relation.
We rely on a classical notion of simulation and perform the same extension as in~\cite{AMHEEMA:2023}, i.e. we start from a simulation relation and add holes and symbolic. The idea is to consider two configurations related by a relation; if one state can do a transition, then the other can also make this transition. 
Like for bisimulation, a simulation relation characterises when two states are related, and this  characterisation is expressed as a predicate on the variables of the two automata.
Simulation defines conditions on a relation $\mathcal{R}$ such that $\mathcal{R}\mpar{s_1, s_2}$ is a predicate (possibly involving variables of the automata) that is true when the state $s_1$ of $A_1$ simulates the state $s_2$ of $A_2$.

@@

 %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).   
The first and second conditions coincide with the natural way to prove inductively that an automaton simulates another by starting with the initial state. The third condition ensures that  simulation  prevents the introduction of deadlocks.
\LUDO{Similarly to bisimulation, the second condition states that, for any transition of the simulating automaton $A_1$, it corresponds to a transition of the automaton $A_2$ that does the same thing and ends up in a similar state. However a family is needed in $A_2$ because of the symbolic nature of transitions, and because depending on the values of the variables, $t_1$ may correspond to different transitions in $A_2$.}
Our definition captures a simple simulation for OA with the same holes that is more expressive than a strict simulation since it matches a transition with a family of transitions. 
For example, with such a relation we are able to check the simulation between two OA that differ by duplicated states, removed duplicated transitions,  reinforced  guards, different variables, etc.
We will show in Section~\ref{sec:prop} that this simulation relation has good properties in terms of transitivity, compositionality, and reflexivity.

@@

In the following definition we add a parameter $H$ which is the set of holes tracked by the simulation relation and adapt the definition by ignoring actions of the holes that are not in $H$.
\LUDO{
Consequently, there is no guarantee related to the actions of the holes outside $H$.
We provide compositionality properties when plugging an automaton inside a hole in $H$ but cannot state anything when plugging an automaton outside $H$. The principle is that any property concerning holes that are not in $H$ should be proven specifically for the considered automaton or the considered composition of automata.
}

\begin{definition}[Hole-tracking simulation]\label{Def:OA-Refinement}
For two OA\\ \(A_1 = \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 = \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a simulation of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation on configurations $\mathcal{R}: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that\footnote{Note that the definition below is identical to the hole equal simulation except $\cap H$ is added in a few places.}:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \models \mathcal{R}\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		 \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\hfill\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \models\\\hspace{5em} \mpar{\mathcal{R}\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{1} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge \mathcal{R}\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2
	\models\! \biggl(\! \mathcal{R}\mpar{s_1, s_2} \!\implies\! \Bigl(\!\! \bigvee_\subbox{t_1 \in \fOT{s_1}}\! \fguard{t_1} \Bigr) \lor \lnot \Bigl(\!\! \bigvee_\subbox{t_2 \in \fOT{s_2}}\! \fguard{t_2} \Bigr) \!\biggr)
\]
\end{definition}
Note that every action of the holes outside \(H\) is unconstrained according to the simulation relation.

>>>>

 \begin{itemize}
 \item   
For any open transition $OT$ in ${T}_1$:
$
     \OTbase{s \OTarrow{\alpha} s'}
         {
           \beta_j^{j\in J'}}{g_{OT}}{\psi_{OT}}
$
 there exists an indexed set of  open transitions $OT_x^{x\in X} \subseteq {T}_2$:
$
%    \left( fresh \ \set{\alpha_i}, \set{b_j}, v_x.\ \
    \OTbase{t \OTarrow{\alpha_x} t_x}
         {
           \beta_{j x}^{j\in J_{x}}}{ g_{OT_x}}{\psi_{OT_x}}   
%         \right)
$
%\end{minipage}
%\hspace{2mm}
%\begin{minipage}{0.35\linewidth}
%\vspace{-2em}
%{	\includegraphics[width=1.02\linewidth]{XFIG/Bisim}}
%\end{minipage}
 such that the following  holds
\begin{multline*}
 \mathcal{R}(s,t) \land g_{OT}\implies\\
 \displaystyle{\bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \land g_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \mathcal{R}\left(s',t_x\right)\psubst{\psi_{OT}\!\uplus\!\psi_{OT_x}}\right)}
\end{multline*}
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.
%\bigskip
%
% $\Pred_{s,t} \land \Pred_{OT}\implies\bigvee_{x\in X} \Pred_x$
%and\\
%%\hspace{1cm}
% $ \forall{x\in X}. \Pred_x \land \Pred_{s,t} \land \Pred_{OT} \Rightarrow
%   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
%     \land \alpha\!=\!\alpha_x \land  
%     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)$
%     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
%     %     \right)$.

@@

Similarly to FH-bisimulation~\cite{AMHEEMA:2023} we are interested  in finding simulation relations between configurations of two OA that contain variables and holes. When dealing with open systems it is common to define simulation in terms of a simulation relation.
We rely on a classical notion of simulation and perform the same extension as in~\cite{AMHEEMA:2023}, i.e., we start from a simulation relation and add holes and symbolic. The idea is to consider two configurations related by a relation; if one state can do a transition, then the other can also make this transition. 
Like for bisimulation, a simulation relation characterises when two states are related, and this  characterisation is expressed as a predicate on the variables of the two automata.
Simulation defines conditions on a relation $\mathcal{R}$ such that $\mathcal{R}\mpar{s_1, s_2}$ is a predicate (possibly involving variables of the automata) that is true when the state $s_1$ of $A_1$ simulates the state $s_2$ of $A_2$.

@@

 %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).   
The first and second conditions coincide with the natural way to prove inductively that an automaton simulates another by starting with the initial state. The third condition ensures that  simulation  prevents the introduction of deadlocks.
Similarly to bisimulation, the second condition states that, for any transition of the simulating automaton $A_1$, it corresponds to a transition of the automaton $A_2$ that does the same thing and ends up in a similar state. However a family is needed in $A_2$ because of the symbolic nature of transitions, and because depending on the values of the variables, $t_1$ may correspond to different transitions in $A_2$.
Our definition captures a simple simulation for OA with the same holes that is more expressive than a strict simulation since it matches a transition with a family of transitions. 
For example, with such a relation we are able to check the simulation between two OA that differ by duplicated states, removed duplicated transitions,  reinforced  guards, different variables, etc.
We will show in Section~\ref{sec:prop} that this simulation relation has good properties in terms of transitivity, compositionality, and reflexivity.

@@

In the following definition we add a parameter $H$ which is the set of holes tracked by the simulation relation and adapt the definition by ignoring actions of the holes that are not in $H$.

Consequently, there is no guarantee related to the actions of the holes outside $H$.
We provide compositionality properties when plugging an automaton inside a hole in $H$ but cannot state anything when plugging an automaton outside $H$. The principle is that any property concerning holes that are not in $H$ should be proven specifically for the considered automaton or the considered composition of automata.

\begin{definition}[Hole-tracking simulation]\label{Def:OA-Refinement}
For two OA\\ \(A_1 = \OA{S_1}{s_{01}}{V_1}{\sigma_{01}}{J_1}{T_1}\) and \(A_2 = \OA{S_2}{s_{02}}{V_2}{\sigma_{02}}{J_2}{T_2}\), \(A_1\) is a simulation of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation on configurations $\mathcal{R}: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that\footnote{Note that the definition below is identical to the hole equal simulation except $\cap H$ is added in a few places.}:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \models \mathcal{R}\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		 \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\hfill\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \models\\\hspace{5em} \mpar{\mathcal{R}\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{1} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge \mathcal{R}\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2
	\models\! \biggl(\! \mathcal{R}\mpar{s_1, s_2} \!\implies\! \Bigl(\!\! \bigvee_\subbox{t_1 \in \fOT{s_1}}\! \fguard{t_1} \Bigr) \lor \lnot \Bigl(\!\! \bigvee_\subbox{t_2 \in \fOT{s_2}}\! \fguard{t_2} \Bigr) \!\biggr)
\]
\end{definition}
Note that every action of the holes outside \(H\) is unconstrained according to the simulation relation.
